Complete Algebraic File Synchronization

### 2015-12-20 2015/def-filesystem.tex
Minor changes, added references

To model this, in $\setv$ we distinguish between 
file contents ($\setf$) and directory contents ($\setd$), that is,
if $\setb = \{\empt\}$ then:
\[ \setv = \setb \cup \setf \cup \setd \]
The tree property can then be defined as
\begin{align*}
\forall n\in\setn: &\FS(n) \neq \empt \\ % TODO alignment problem
&\quad\wedge \parent(n) \neq \topnode \\
&\Rightarrow \FS(\parent(n)) \in \setd 
\end{align*}
This means that as we move down from the root of a tree of nodes,
the types of values we encounter in the filesystem can only change according to the
transition diagram in \cref{fig_transition}.

>>>>

To model this, in $\setv$ we distinguish between 
file contents ($\setf$) and directory contents ($\setd$), that is,
if $\setb = \{\empt\}$ then:
\[ \setv = \setb \cup \setf \cup \setd \]
The tree property can then be defined as
\begin{align*}
\forall n\in\setn: &\FS(n) \neq \empt \\ % TODO alignment problem
&\quad\wedge \parent(n) \neq \topnode \\
&\Rightarrow \FS(\parent(n)) \in \setd 
\end{align*}
This means that as we move down from the root of a tree of nodes,
the types of values we encounter in the filesystem can only change according to the
transition diagram in \cref{fig_transition}, from directories ($\cchard$) to files ($\ccharf$)
to empty nodes ($\ccharb$).

### 2015-12-20 2015/intro.tex
Minor changes, added references

A central problem of command-based synchronizers during both update detection and reconciliation
is ordering.
If update detection is based on comparing the original state of the filesystem
to its new state, then we can easily collect a set of updates,
but we need to order these in a way that they could be applied in
without causing an error 
(for example, a directory needs to be created first before a file can be created under it).
Similarly, during reconciliation when we consider all the updates that
have been applied to the different replicas, we need to find a way of
merging these and creating a global ordering that we can apply
to the filesystems.
In fact, an insight in \cite{NREC} is that
if the ordering changes the effect of two commands,
then under certain circumstances they are not compatible and will give rise to conflicts.
Accordingly, in this paper command-pairs and their properties
play a special role.

In \cref{sec_def} we define a model of filesystems and a set of commands
that we will use to describe updates and modifications.
This is followed by investigating the properties and behaviors
of command pairs in \cref{section_axioms}.
\Cref{sec_update} describes update detection, and it is
in \cref{ordering} that we describe how commands can be ordered
and prove that under some conditions are valid orders
have the same effect.

Finally, \cref{sec_rec} defines our reconciliation algorithm
and when two updates from different replicas give rise to a conflict.
We then proceed to prove that the output of this algorithm
is sound in the sense that 
it can be applied to the replicas without causing an error,
and that the described algorithm provides a complete solution
inasmuch as no further updates can be applied under any circumstances.
In order to be able to do this, we introduce
the concept of {\emph conditional operation} of commands,
and show that it has a number of highly convenient properties.

The approach in the present paper offers an improvement
over results presented in \cite{NREC} in multiple ways.
We introduce a new set of commands that is symmetric and
captures more information about the updates.
This has the effect that reasoning about commands
becomes simpler as there are fewer edge cases,
and more powerful as with the additional information
content predictions made by the reasoning are more accurate.
In fact, the new command set not only simplifies
our reconciliation algorithm, but also makes it complete.

The previous work also lacked proofs that the update detection
and reconciliation algorithms it presented work as intended,
which we provide in the current paper.
While the results are intuitive, providing rigorous proofs
is far from trivial.
During the process, we define a number of useful concepts
and show their relationships and the properties they possess.
In our view these construct an algebraic model
that is worthy of interest and further research on its own.

>>>>

A central problem of command-based synchronizers during both update detection and reconciliation
is ordering.
If update detection is based on comparing the original state of the filesystem
to its new state, then we can easily collect a set of updates,
but we need to order these in a way that they could be applied
to the filesystems
without causing an error 
(for example, a directory needs to be created first before a file can be created under it).
Similarly, during reconciliation when we consider all the updates that
have been applied to the different replicas, we need to find a way of
merging these and creating a global ordering that we can apply
to the filesystems.
In fact, an insight in \cite{NREC} is that
if the ordering changes the effect of two commands,
then under certain circumstances they are not compatible and will give rise to conflicts.
Accordingly, in this paper command-pairs and their properties
play a special role.
Other research also shows the essential nature of ordering:
in IceCube multiple orders are tested to find an acceptable one \cite{KRSD},
or in Bayou reconciling updates happens by redoing them in an acceptable order \cite{TTPDSH}.

In \cref{sec_def} we define a model of filesystems and a set of commands
that we will use to describe updates and modifications.
This is followed by investigating the properties and behaviors
of command pairs in \cref{section_axioms}.
\Cref{sec_update} describes update detection, and it is
in \cref{ordering} that we describe how commands can be ordered
and prove that under some conditions all valid orders
have the same effect.

Finally, \cref{sec_rec} defines our reconciliation 
and conflict detection algorithms.
We then proceed to prove that the output of reconciliation
is sound in the sense that 
it can be applied to the replicas without causing an error,
and that the described algorithm provides a complete solution
inasmuch as no further updates can be applied under any circumstances.
In order to be able to do this, we introduce
the concept of \emph{conditional operation} of sequences of commands,
and show that it has a number of highly convenient properties.

The approach in the present paper offers an improvement
over results presented in \cite{NREC} in multiple ways.
We introduce a new set of commands that is symmetric and
captures more information about the updates.
This has the effect that reasoning about commands
becomes simpler as there are fewer edge cases,
and more powerful as 
predictions made by the reasoning are more accurate
due to the additional information content.
In fact, the new command set not only simplifies
our reconciliation algorithm, but also makes it complete.

The previous work also lacked proofs that the update detection
and reconciliation algorithms it presented work as intended,
which we provide in the current paper.
While the results are intuitive, providing rigorous proofs
is far from trivial.
During the process, we define a number of useful concepts
and show their relationships and the properties they possess.
In our view these construct a special algebraic model
that is worthy of interest and of further research on its own.

### 2015-12-20 2015/works-inverse.tex
Minor changes, added references

For the proposition
\[ \worksc{B\cc \seqset{A}}{\seqset{S}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{S}} \]
the rest of the proof is illustrated by \cref{fig_invmove}
where Dom($x$), the domain of $x$, represents the set of filesystems
the sequence $x$ does not break, or neither of the sequences in the set $x$ breaks.

>>>>

For the proposition
\[ \worksc{B\cc \seqset{A}}{\seqset{S}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{S}} \]
the rest of the proof is illustrated by \cref{fig_invmove}
where $\textrm{Dom}(x)$, the domain of $x$, represents the set of filesystems
the sequence $x$ does not break, or neither of the sequences in the set $x$ breaks.
